---
title: Programming Languages Visualisation
author: Rosa Pham
output:
  html_notebook:
    toc: true
---

```{r}
knitr::opts_chunk$set(warning = FALSE) 
```

### Import libraries

```{r}
library(ggplot2)
library(dplyr) 
library(data.table)
library(forcats)
library(viridis)
library(hrbrthemes)
library(ggpubr)
library(wordcloud)
```

### Data Processing
- Load csv file

```{r include=TRUE}
data <- read.csv('./pldb.csv')
head(data)
```

- Check the number of observations in the data

```{r}
nrow(data)
```

- Check the number of NA values in each column

```{r}
na_count <- data %>% summarise_all(funs(sum(length(which(is.na(.)))))) %>% data.frame() %>% melt(variable.name = "variables", value.name = "na_count")
na_count
```

- Check the number of empty string in each column

```{r}
empty_count <- data %>% summarise_all(funs(sum(. == ""))) %>% data.frame() %>% melt(variable.name = "variables", value.name = "empty_count")
empty_count
```

- Select the neccessary columns for visualisation

```{r}
visual_data <- data %>% select(c('title', 'appeared', 'type', 'languageRank', 'numberOfUsers', 
                                   'numberOfJobs', 'lastActivity', 'factCount', 'bookCount', 
                                   'wikipedia.dailyPageViews', 'country'))
head(visual_data)
```

- Only get observations with type as programming language
```{r}
visual_data <- visual_data[visual_data$type %in% c('pl', 'queryLanguage', 'stylesheetLanguage', 'contractLanguage'),]
head(visual_data)
```

- Check the number of observations after filtering programming language type

```{r}
nrow(visual_data)
```

### Data Visualisaton
##### 1. Get the top 10 programming languages ordered by numberOfUsers

```{r}
visual_data %>% arrange(desc(numberOfUsers)) %>% head(10) %>%
  mutate(title = fct_reorder(title, numberOfUsers)) %>%
  ggplot( aes(x=title, y=numberOfUsers)) +
  geom_bar(stat="identity", fill="royalblue") +
  scale_y_continuous(name="The number of Users", labels = scales::comma) +
  coord_flip() +
  xlab("") +
  ggtitle("The top 10 Programming Languages by Users") +
  theme_bw() +
  theme(plot.title = element_text(size=14, face="bold", margin = margin(0, 0, 15, 0)))
```

##### 2. What are the top programming languages based on their ranking?

```{r}
incremental_list <- c(50, 15, 13, 10, 7, 5, 3, 1)
wordcloud_data <- visual_data %>% select('title', 'languageRank') %>% head(100)
wordcloud_data <- wordcloud_data %>% mutate(ranking = rank(-wordcloud_data$languageRank))
wordcloud_data <- wordcloud_data %>% arrange(desc(ranking)) %>% select('title', 'ranking')
for(i in 1:length(incremental_list)) {
  wordcloud_data[i, 2] <- wordcloud_data[i, 2] + incremental_list[i]
}

set.seed(1234)
wordcloud(words = wordcloud_data$title, freq = wordcloud_data$ranking, random.order=FALSE, 
          min.freq = 1, max.words=300, rot.per=0.30, colors=brewer.pal(8, "Dark2"), scale=c(3.5, 0.25))
```

##### 3. Which programming language is stated the most in job requirements
- Get the top 3 programming languages for each types

```{r}
job_type_data <- visual_data %>% 
  arrange(desc(numberOfJobs)) %>% 
  group_by(type) %>% 
  slice(1:3)
head(job_type_data)
```

- Remove types have the numberOfJobs at 0 for any programming language

```{r}
job_type_data <- job_type_data[!!ave(job_type_data$numberOfJobs, job_type_data$type, FUN=function(i) !any(i == 0)),]
types <- c('pl', 'queryLanguage')
```

```{r}
three_types_data <- job_type_data[job_type_data$type %in% types,]
job_type_data$title <- as.character(job_type_data$title)
job_type_data$title <- factor(job_type_data$title, levels=unique(job_type_data$title))
```

```{r}
ggplot(job_type_data, aes(x=title, y = numberOfJobs, fill = type)) + 
  geom_bar(stat = "identity", position = position_dodge()) + 
  scale_fill_discrete(labels=c('Programming Language', 'Query Language')) +
  labs(x = "", y = "Number of Jobs", fill=NULL) + 
  ggtitle("The top 3 languages were stated in job requirements") +
  theme_bw()
```

##### 4. What are the total number of programming languages appeared from 1952 to 2022

```{r}
data_group_year <- visual_data %>% mutate(year = cut(appeared, breaks = c(1938, 1943, 1948, 1953, 1958, 1963, 1968, 1973, 1978, 
                                                                   1983, 1988, 1993, 1998, 2003, 2008, 2013, 2018, 2023), 
                                              labels = c(1943, 1948, 1953, 1958, 1963, 1968, 1973, 1978, 
                                                         1983, 1988, 1993, 1998, 2003, 2008, 2013, 2018, 2023))) %>% 
  group_by(year) %>% summarise(count = n()) %>% na.omit()

data_group_year <- visual_data %>% mutate(year = cut(appeared, breaks = c(1937, 1942, 1947, 1952, 1957, 1962, 1967, 1972, 1977, 
                                                                            1982, 1987, 1992, 1997, 2002, 2007, 2012, 2017, 2022), 
                                                       labels = c(1942, 1947, 1952, 1957, 1962, 1967, 1972, 1977, 
                                                                  1982, 1987, 1992, 1997, 2002, 2007, 2012, 2017, 2022))) %>% 
  group_by(year) %>% summarise(count = n()) %>% na.omit()

ggplot(data_group_year, aes(x=year, y=count, group = 1)) +
  geom_line(color="grey") +
  geom_point(shape=21, color="black", fill="#69b3a2", size=4)  +
  theme_ipsum() +
  ggtitle("The number of languages created from 1952 to 2022")
```

##### 5. Which country is the most programming languages created until 2023

```{r}
country_data <- visual_data %>% group_by(country) %>% summarise(count = n()) %>% filter(count > 1, country != '')
map.world <- map_data("world")
as.factor(country_data$country) %>% levels()
country_data$country <- recode(country_data$country
                                   ,'United States' = 'USA'
                                   ,'United Kingdom' = 'UK'
)

map.world_joined <- left_join(map.world, country_data, by = c('region' = 'country'))
map.world_joined <- map.world_joined %>% mutate(fill_flg = ifelse(is.na(rank),F,T))

df.country_points <- data.frame(country = c("USA","UK"),stringsAsFactors = F)
glimpse(df.country_points)

geocode.country_points <- geocode(df.country_points$country)

ggplot() +
  geom_polygon(data = map.world_joined, aes(x = long, y = lat, group = group, fill = fill_flg)) +
  geom_point(data = df.country_points, aes(x = lon, y = lat), color = "#e60000") +
  scale_fill_manual(values = c("#CCCCCC","#e60000")) +
  labs(title = 'Countries with highest "talent competitiveness"'
       ,subtitle = "source: INSEAD, https://www.insead.edu/news/2017-global-talent-competitiveness-index-davos") +
  theme(text = element_text(family = "Gill Sans", color = "#FFFFFF")
        ,panel.background = element_rect(fill = "#444444")
        ,plot.background = element_rect(fill = "#444444")
        ,panel.grid = element_blank()
        ,plot.title = element_text(size = 30)
        ,plot.subtitle = element_text(size = 10)
        ,axis.text = element_blank()
        ,axis.title = element_blank()
        ,axis.ticks = element_blank()
        ,legend.position = "none"
  )
```

##### 6. What programming language has the greatest number of published books across their types?

```{r}
book_type_data <- visual_data %>% 
  arrange(desc(bookCount)) %>% 
  group_by(type) %>% 
  slice(1:5) %>% filter(bookCount > 0)

typenames <- c("contract language", "programming language", "query language", "stylesheet language")
names(typenames) <- c("contractLanguage", "pl", "queryLanguage", "stylesheetLanguage")

ggplot(book_type_data, aes(x=fct_reorder(title, bookCount), y=bookCount, fill=title)) +
  facet_grid(~type, scale="free_x", space = "free_x", labeller = labeller(type = typenames)) +
  geom_bar(stat = "identity") +
  geom_text(aes(x=fct_reorder(title, bookCount), y=bookCount, label = bookCount), vjust = -0.5) +
  labs(x = "",
       y = "Number of Books") +
  theme(legend.position = "none") 
```

##### 7. Languages were created in USA

```{r}
lang_country_data <- visual_data[visual_data$country %in% c("United States"),] %>% 
  head(10) %>% mutate(ranking = languageRank + 1) %>% select(c("title", "ranking"))

ggplot(lang_country_data, aes(x = reorder(title, -ranking), y = ranking)) +
  geom_segment(aes(x = reorder(title, -ranking),
                   xend = reorder(title, -ranking),
                   y = 0, yend = ranking),
               color = "gray", lwd = 1) +
  geom_point(size = 5.5, pch = 21, bg = 4, col = 1) +
  geom_text(aes(label = ranking), color = "white", size = 3) +
  xlab("") +
  ylab("Ranking") +
  coord_flip() +
  theme_minimal()
```

##### 8. The top page views per day for languages

```{r}
pageviews_data <- visual_data %>% 
  subset(title != "Regular Expressions" & title != "GAME") %>%
  arrange(desc(wikipedia.dailyPageViews)) %>% 
  head(20) %>%
  select(c("title", "wikipedia.dailyPageViews"))

pageviews_data$title[pageviews_data$title == 'Arduino Programming Language'] <- 'Arduino'
pageviews_data$label <- paste0(pageviews_data$title, "-", pageviews_data$wikipedia.dailyPageViews)

# Set a number of 'empty bar' to add at the end of each group
pageviews_data$id <- seq(1, nrow(pageviews_data))

# Get the name and the y position of each label
label_data <- pageviews_data
number_of_bar <- nrow(label_data)
angle <- 90 - 360 * (label_data$id-0.5) /number_of_bar     # I substract 0.5 because the letter must have the angle of the center of the bars. Not extreme right(1) or extreme left (0)
label_data$hjust <- ifelse( angle < -90, 1, 0)
label_data$angle <- ifelse(angle < -90, angle+180, angle)

# Make the plot
ggplot(pageviews_data, aes(x=as.factor(id), y=wikipedia.dailyPageViews)) +       # Note that id is a factor. If x is numeric, there is some space between the first bar
  geom_bar(stat="identity", fill=alpha("skyblue", 1)) +
  ylim(-2000,8000) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = unit(rep(-1,4), "cm") 
  ) +
  coord_polar(start = 0) + 
  geom_text(data=label_data, aes(x=id, y=wikipedia.dailyPageViews+10, label=title, hjust=hjust), color="black", fontface="bold",alpha=0.6, size=2.5, angle= label_data$angle, inherit.aes = FALSE ) 


ggplot(pageviews_data, aes(x = as.factor(id),
                 
                 # y-axis numerical parameter
                 y = wikipedia.dailyPageViews)) +     
  
  # the bar height will represent
  # the actual value of the data
  geom_bar(stat = "identity",
           
           # define bar color
           fill=alpha("green", 0.5)) +
  
  # define size of inner circle
  # and the size of the bar
  ylim(-1000,8000) +
  # define the polar coordinate
  coord_polar(start = 0) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.title = element_blank(),
    plot.margin = unit(rep(-1,4), "cm") 
  ) +
  # add labels
  geom_text(data = label_data,
            aes(x = id, y = wikipedia.dailyPageViews+10,
                
                # label alignment
                label = title, hjust=hjust),
            color = "black", fontface="bold",
            alpha = 0.8, size = 2.5,
            angle = label_data$angle,
            inherit.aes = FALSE )
```


```{r}
rmarkdown::render("program_notebook.Rmd", "all")
```


